### 2012-03-01 2.1.3.tex
Drastic restructuring and some writing on chapters 2-3.

I assume that the [j] onglide in words such as \emph{ass}[j]\emph{ume} is present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}). The primary motivation for this assumption is the observation that presence or absence of this glide is contrastive (e.g. \emph{coup} $\sim$ \emph{queue}, \emph{booty} $\sim$ \emph{beauty}). I further assume that the front glide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}. However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm]. This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear. There is a great deal of external support for this position as well. The [ju\lm] in words like \emph{spew} may pattern together
in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

>>>>

I assume that the [j] onglide in words such as \emph{ass}[j]\emph{ume} is present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}). The primary motivation for this assumption is the observation that presence or absence of this glide is contrastive (e.g. \emph{coup} $\sim$ \emph{queue}, \emph{booty} $\sim$ \emph{beauty}). I further assume that the front glide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}. However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm], and speakers judge other following vowels to be anomalous in nonce words \citep{Moreland2009}. This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear. There is a great deal of external support for this position as well. The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

### 2012-03-01 2.2.1.tex
Drastic restructuring and some writing on chapters 2-3.

\subsubsection{Statistical technique}

@@

\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
\chi^2$ \xe

@@

English has a process of voice assimilation which is found in two inflectional suffixes. The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg. present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S. \citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}

@@

\ex \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178): \\
\xymatrix@C=6pt@R=12pt{
\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}}
\label{ovar}
\xe

@@

\ex \textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85): \\
\label{CNPA}
\xymatrix@C=6pt{
                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} }
\xe

@@

There is some evidence that even young infants show a preference for nonce words that conform to this generalization. \citet{Jusczyk2003}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}). \citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''. In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.

@@

\ex \textsc{Degemination} (\emph{SPE}:253): \\
\label{degem}
%\xymatrix@C=6pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
%\textsc{C}\ar@{-}[d]       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}\ar@{-}[d] \\
%*+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} & & & & & *+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} \\ \\
%}
\xymatrix@C=6pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
\textsc{C}_1       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}_2 }
\\ 
Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
\xe

>>>>

\subsubsection{Statistical technique}
\label{stattech}

@@

\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
χ^2$ \xe

@@

English has a process of voice assimilation which is found in two inflectional suffixes. The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg. present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S. \citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}

@@

\ex \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178): \\
\xymatrix@R=12pt{
\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}}
\label{ovar} \xe

@@

\ex \textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85): \\
\label{CNPA}
\xymatrix@R=12pt{
                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} }
\xe

@@

There is some evidence that even young infants show a preference for nonce words that conform to this generalization. %\citet{Jusczyk2003}
\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}). \citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''. In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.

@@

\ex \textsc{Degemination} (\emph{SPE}:253): \\
\label{degem}
%\xymatrix@C=6pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
%\textsc{C}\ar@{-}[d]       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}\ar@{-}[d] \\
%*+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} & & & & & *+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} \\ \\
%}
\xymatrix@R=12pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
\textsc{C}_1       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}_2 }
\\ 
Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
\xe

### 2012-03-01 2.2.3.tex
Drastic restructuring and some writing on chapters 2-3.

That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation. Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}. The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.

>>>>

That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation. Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also word \citep{Baroni2009} and phoneme \citep{Daland2011a} n-grams, and tokens in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}. The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.

### 2012-03-01 3.1.1.tex
Drastic restructuring and some writing on chapters 2-3.

% 3.1.1: Alternation vs. phonotactic

\subsubsection{Roundness harmony}

\subsubsection{Labial attraction}

\citet{Zuraw2000}
\citet{Clements1982}
\citet{Inkelas1997}

>>>>

% 3.1.1: Backness harmony

\subsubsection{Phonological description}

One of the most salient aspects of Turkish phonology is the presence of vowel harmony. \citeauthor{Lees1966a} (\citeyear[][284]{Lees1966b}, \citeyear[][35]{Lees1966a}) proposes a number of phonological rules to describe the harmony processes that act on roots and regularly spread rightward to suffixes (Turkish has no prefixes); one such rule consists of rightward spreading of the feature [\textsc{Back}].

\ex \textsc{Backness Harmony} \citep[e.g.,][229]{Clements1982}: \\
\xymatrix@R=12pt{
\txt{[α \textsc{Back]}}\ar@{-}[d]\ar@{..}[drr] &             \\
\txt{V}                                        & \txt{C$_0$} & \txt{V}}
\xe

%\citet{Clements1982}
%\citet{Inkelas1997}

This rule accounts for much of the suffix allomorphy found in Turkish. The dat.sg. (and the vowel of the gen.sg.) is always a [$+$\textsc{High}] vowel, and in general, it agrees with the final stem vowel for backness and roundness. Similarly, the vowel of the nom.pl. is [$-$\textsc{High}, $-$\textsc{Round}], and takes its backness from the stem-final vowel. Similar patterns obtain for many other Turkish suffixes.

\ex Turkish nominal suffix allomorphy: \vspace{6pt} \\
\begin{tabular}{l l l l l l l}
   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.} & \emph{nom.pl.} \\
a. & ip             & ipi            & ipin           & ipler          & `rope' & (C\&S:216) \\
   & el             & eli            & elin           & eller          & `hand'    \\
   & kız            & kızı           & kızın          & kilzar         & `girl'    \\
   & sap            & sapı           & sapın          & saplar         & `stalk'   \\
   & yüz            & yüzü           & yüzün          & yüzler         & `face'    \\
   & köy            & köyü           & köyün          & köyler         & `village' \\
   & pul            & pulu           & pulun          & pullar         & `stamp'   \\
   & son            & sonu           & sonun          & sonlar         & `end'     \\
b. & neden          & nedeni         & nedenin        & nedenler       & `reason'  & (TELL) \\
   & boğaz          & boğazı         & boğazın        & boğazlar       & `throat'  \\
   & kiler          & kileri         & kilerin        & kilerler       & `pantry'  \\
   & pelür          & pelürü         & pelürün        & pelürler       & `tissue paper' \\
   & sapık          & sapığı         & sapığın        & sapıklar       & `pervert' \\
\end{tabular} \xe

\noindent
If the vowel of the nom.pl. suffix is specified as [$-$\textsc{High}, $-$\textsc{Round}], then the \emph{a} $\sim$ \emph{e} pattern follows naturally from \textsc{Backness Harmony}. If the dat.sg. suffix is a vowel specified only as [$+$\textsc{High}], then \textsc{Backness Harmony}, along with the rightward spreading of \textsc{Round}, will produce the correct results.

There are a few complications, however. First, while most polysyllabic roots conform to \textsc{Backness Harmony} (\lastx b), not all do. In this case, suffix vowels generally harmonize with the final root vowel (\nextx a). Finally, here is a small class of nouns which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}] (\nextx b).

\ex Turkish nominal suffix allomorphy: \vspace{6pt} \\
\begin{tabular}{l l l l l l l}
   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.} & \emph{nom.pl.} \\
a. & mezar          & mezarı         & mezarın        & mezarlar       & `grave' & (TELL) \\
   & model          & modeli         & modelin        & modeller       & `model' \\
   & silah          & silahı         & silahın        & silahlar       & `weapon'     \\
   & memur          & memuru         & memurun        & memurlar       & `bureaucrat' \\
   & sabun          & sabunu         & sabunun        & sabunlar       & `soap'       \\
b. & saat           & saati          & saatin         & saatler        & `hour, clock' \\
   & harf           & harfi          & harfin         & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
   & etol           & etolü          & etolün         & etoller        & `fur stole' \\
\end{tabular} \xe

While it is uncontroversial that disharmonic suffixes (\lastx b) constitute no more than sporadic exceptions to \textsc{Backness Harmony}, the status of root disharmony (\lastx a) has been the subject of much debate.

1. \citet[][289]{Anderson1974} 
2. \citet[][]{Kiparsky1981} makes a similar objection regarding Finnish vowel harmony.
3. \citet{Clements1982}
4. \citet{Inkelas1997}

have identified (\lastx b) as sporadic exceptions to 
A number of objections can be made for

\subsubsection{Lexical statistics}

\citet{TELL}
%\citet{Inkelas2001}

\ex Lexical effects of \textsc{Obstruent Voice Assimilation}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
violating  &  6 & 264 & 0.022 \\
\bottomrule
\end{tabular} \xe

\subsubsection{External evidence}

The data reviewed thus far leaves open the possibility that speakers do not internalize the generalization that roots conform to \textsc{Backness Harmony}. One piece of evidence against this hypothesis comes from a Turkic language game \citep{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it corresponds to a morphological process native to the related langauge Tuvan, and is quickly taught to even young Turkish speakers. The process consists of reduplication of the base with the base-initial vowel replaced by \emph{a} or \emph{u}, which conveys a sense of ``vagueness'' or ``jocularity''. Reduplication interacts with root harmony in both Tuvan and Turkish. When the base is harmonically [$-$\textsc{Back}], the non-initial vowels of the reduplicant do not surface faithfully: they surface as [$+$\textsc{Back}] (\nextx a).

\ex Turkish reduplication game \citep[][231]{Harrison2001}: \\
\begin{tabular}{l l l l}
a. & kibrit & kibrit-kabrıt & `match'    \\
   & bütün  & bütün-batın   & `whole'    \\
b. & mali   & mali-muli     & `Mali'     \\
   & butik  & butik-batik   & `boutique' \\
\end{tabular} \xe

\noindent
Interestingly, reharmonization is absent in disharmonic bases (\lastx b). This is not an isolated fact about Turkic, as \citeauthor{Harrison2001} note similar outcomes in an unrelated Finnish language game \citep{Campbell1986}. The simplest explanation of this contrast is to treat \textsc{Backness Harmony} as a feature-filling process, as proposed by \citet{Clements1982} and \citet{Inkelas1997}. In harmonic roots, the non-initial vowels are underspecified for \textsc{Back} and are spelled out by \textsc{Backness Harmony}. In reduplication, this produces reharmonization. In disharmonic roots, the vowels are fully specified and do not meet the structural description of \textsc{Backness Harmony}.

A number of studies have found that speakers of Finnish use disharmony as a clue to segment running speech.\footnote{Thanks to Charles Yang for pointing out the relevance of these studies to me.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which are harmonic for the feature \textsc{Back}. These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word. Speakers are quicker to press the button when the prefix and disyllabic word disagree for \textsc{Back}. This is taken to indicate that speakers are attuned to the fact that disharmonic transistions are good predictors of word boundaries. On the other hand, if this also indicates that speakers recognize that harmonic transistions are more likely to be root-internal, then harmony is active not just where it is triggered by alternations (i.e., in Finnish affixes) but also in roots.

\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when it preceded by the disharmonic pseudo-prefix \emph{gölü-} than when it is prefixed with harmonic nonce prefix \emph{golu-}. The effect of harmony does not obtain for speakers of French, which lacks vowel harmony. As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.

It seems that root-internal harmony is not only present in adults, but in fact is learned very early. The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}. Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic. At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm. Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix. This preference is not observed in 9-month-old infants learning German, which lacks vowel harmony. Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants do not.

%We might suggest methods not based only on statistical significance, but rather substantive computational properties of the linguistic objects in question \citep[e.g.,][]{Yang2005a}.

### 2012-03-01 3.1.2.tex
Drastic restructuring and some writing on chapters 2-3.

% 3.1.2: Lexical statistics

This is essentially the hypothesis pursued in the previous chapter.

\citet{TELL}
\citet{Inkelas2001}

%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.

>>>>

% 3.1.2: Labial attraction

I am using the dotted association line in a deliberately ambiguous fashion in this chapter. The autosegmental structures represent formalizations of the constraints, but I intend to remain ambiguous as to their ontology and do not intend them to be interpreted as derivational rules. These questions are taken up more carefully below.

\citet{Clements1982}

\citet{Zimmer1969}, \citet{Zuraw2000}, and \citet{Inkelas2001} all suggest

\subsubsection{Phonological description}

\citet[][36]{Lees1966a} describes \textsc{Labial Attraction} as a process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified''. This description is transalted into an autosegmental rule below.

\ex \textsc{Labial Attraction} (after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): \\
\xymatrix@R=12pt{
\txt{[$-$\textsc{Round}]}\ar@{-}[d] &                            & \txt{[$+$\textsc{Round}]}\ar@{..}[d] \\
\txt{V}\ar@{-}[d]                    & \txt{C}\ar@{-}[d]          & \txt{V}\ar@{-}[d] \\
\txt{[$-$\textsc{High}, $+$\textsc{Back}]}             & \txt{[$+$\textsc{Labial}]} & \txt{[$+$\textsc{High}, $+$\textsc{Back}]}}
\xe

\noindent
A few examples of words conforming to this generalization are \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', and \emph{samur} `sable' \citep[][285]{Lees1966b}; \citeauthor{Lees1966b} also notes the existence of exceptions like \emph{tavır} `mode', but writes they are ``surprisingly rare'' (ibid., 286); \citet[][225]{Clements1982} lists a number of additional exceptions.

\citeauthor{Clements1982} propose that there is an additional type of exception to \textsc{Labial Attraction}.

\begin{quotation}
Even more decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in rotos of the form /\ldots~aCu~\ldots/. \citep[][225]{Clements1982}
\end{quotation}

\noindent
\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1996b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.

There is reason to believe that \textsc{Labial Attraction} is at best a static constraint: it never applies across morpheme boundaries. If, contrary to fact, there was a \textsc{Labial Attraction} alternation, then one would expect, for example, that the gen.sg. of \emph{sap} `stalk' would be *\emph{sapun} instead of the observed \emph{sapın}. Additional potential evidence for the static nature of \textsc{Labial Attraction} is provided by \citeauthor{Clements1982}, who note a class of bisyllabic words in which the second vowel, always [$+$\textsc{High}], alternates with zero.

\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
\begin{tabular}{l l l l}
   & nom.sg. & 3.poss. \\
a. & fikir   & fikri & `idea' \\ 
   & hüküm   & hükmü & `judgement' \\ 
%   & filim   & filmi & `film' & \citep[][178]{Inkelas2001} \\
%   & kojun   & kojun & `bosom' \\
b. & vakit  & vakti  & `time' \\
   & rahim  & rahmi  & `womb' \\
\end{tabular}
\xe

\noindent 
The vowel is present in the nom.sg. form and absent in the 3.poss. form. The examples in (\lastx b) shows that it need not harmonize with the preceding vowel
\citeauthor{Inkelas2001} (ibid, 178) note that \textsc{Labial Attraction} does not affect this vowel (e.g., \emph{sabır}/\emph{sabrı} `patience'), citing this as evidence against \textsc{Labial Attraction}. However, it is not a foregone conclusion that \textsc{Labial Attraction} should apply to the ``unstable'' vowel. Since it is as a morpheme structure constraint, this vowel must be underlying present and deleted in the 3.poss.; \citeauthor{Clements1982} prefer the converse analysis, in which epenthesis applies in the nom.sg., though they note that either approach requires lexically arbitrary diacritics. I conclude that unstable vowels provide no evidence for or against \textsc{Labial Attraction}.

\begin{quotation}
Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon. It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic. \citep[][196]{Inkelas2001}
\end{quotation}

\begin{quotation}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[][196, emphasis in original]{Inkelas2001}
\end{quotation}

\citet{Inkelas1997}
\citet{Inkelas2001}
\citet{TELL}

\subsubsection{Lexical statistics}

%##TELL: 31,236
%#Lexicon: 25,164

\ex Lexical effects of \textsc{Labial Attraction} \citep[][186]{Inkelas2001}: \vspace{6pt} \\
\begin{tabular}{l r r r | r r r}
\toprule
Corpus  & aPu & aPı & \% & aTu & aTı & $p$-value   \\ % & corpus size
\midrule
%TELL    & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
%TELL    & 77  & 30 & 72.6 & 115 & 356 & ?\e{-20} \\
%Lexicon & 108 & 57 & 65.5 & 155 & 533 & 6.6\e{-25} \\
\bottomrule
\end{tabular} \xe

BROAD labial attraction stats:

TELL full:
585 aPu (favored)
230 aPI (disfavored)
1,137 aTu
1,461 aTI
71.8\% labial attraction
p = 3.4\e{-45}

Lexicon:
1,175 aPu (favored)
446 aPI (disfavored)
1,939 aTu
2,878 aTI
72.5\% labial attraction
p = 3.1\e{-114}

STUPID attraction stats:

TELL full:
267 aPu (favored)
546 aPI (disfavored)
346 aTu
327 aTI
67.1\% attraction
p = 5.2\e{-13}

Lexicon:
357 aPu (favored)
874 aPI (disfavored)
488 aTu
396 aTI
71% attraction
p = 9.7\e{-34}

zimmer: 319-320 
inkelas et al.: 196

\subsubsection{External evidence}

Beyond the experimental results of \citet{Zimmer1969} reviewed in the next section, there is no external evidence for or against \textsc{Labial Attraction}.

%% HERE THERE BE DRAGONS
%\ex Examples of \textsc{Labial Attraction} \citep[][285]{Lees1966b}: \\
%\begin{tabular}{l l l}
%a. & çapul &  `raid \\
%b. & sabur & `patience' \\
%c. & şaful & `wooden honey tub' \\
%d. & avuç  & `palm of hand' \\
%   & havuz & `hollow'       \\
%e. & samur & sable'         \\
%   & camus & `buffalo'      \\
%\end{tabular}
%\xe

### 2012-03-01 3.1.tex
Drastic restructuring and some writing on chapters 2-3.

>>>>

I assume without discussion the following binary feature specifications for the vowel phonemes of Turkish.

\ex Turkish vowel features: \vspace{6pt} \\
\begin{tabular}{c | c c c c}
                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
\end{tabular}
\xe

%\citep[][289]{Anderson1975}
%\citep{Clements1982}
%\citet{Inkelas1997}
%\citet{Harrison2001}

\subsection{Backness harmony}  \input{3.1.1}
\subsection{Labial attraction} \input{3.1.2}

### 2012-03-01 3.2.2.tex
Drastic restructuring and some writing on chapters 2-3.

>>>>

There is some disagreement in the literature about just what \citet{Zimmer1969} finds regarding \textsc{Labial Attraction}.

\citet{Zimmer1969}
(see also \citet[][4]{Zuraw2000})
claims however, Zimmer's evidence suggests that the speakers did internalize a version of the labial attraction generalization.

\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}

\begin{quotation}
Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research. \citep[][196, emphasis in original]{Inkelas2001}
\end{quotation}

``These results support our view that there is no rule of Labial Attraction'' \citep[][225]{Clements1982}.

since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.

\citet{Ito1993} argue that the data shows that it holds of native words only
